Real-space computation of $E$/$B$-mode maps I: Formalism and Compact Kernels

### 2018-06-11 articles/formalism/article.tex
Added some reference and some detail to kernel paintbrush section.
Split kernel image and moved MG/MB to right after they are introduced.
Working some on introduction
significant updates to discussions and polarized filaments sections

\usepackage[latin1]{inputenc}
\usepackage{amsmath, amsfonts, amssymb, hyperref, graphicx, multirow, subfigure, bbold}
\hypersetup{colorlinks=True,citecolor=blue}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage[normalem]{ulem}
\usepackage{caption}
\renewcommand{\thesubfigure}{(\roman{subfigure})}
%\renewcommand{\thesubfigure}{}

\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}
\newcommand{\beqry}{\begin{eqnarray}}
\newcommand{\eeqry}{\end{eqnarray}}
\newcommand{\beqrys}{\begin{subequations}\begin{eqnarray}}
\newcommand{\eeqrys}{\end{eqnarray}\end{subequations}}
\newcommand{\fqu}{\begin{bmatrix} Q \\ U \end{bmatrix}}
\newcommand{\xpm}{\begin{bmatrix} _{+2}X \\ _{-2}X \end{bmatrix}}
\newcommand{\equ}{\begin{bmatrix} Q \\ U \end{bmatrix}_{E}}
\newcommand{\bqu}{\begin{bmatrix} Q \\ U \end{bmatrix}_{B}}
\newcommand{\eblm}{\begin{bmatrix} a^E \\ a^{B} \end{bmatrix}}
\newcommand{\qutox}{\begin{bmatrix} \mathbb{1} & i\mathbb{1}  \\  \mathbb{1} & -i\mathbb{1}   \end{bmatrix}}
\newcommand{\qutoxd}{\begin{bmatrix} \mathbb{1} & \mathbb{1}  \\  -i\mathbb{1} & i\mathbb{1}   \end{bmatrix}}
\newcommand{\ymat}{\begin{bmatrix} _{+2}Y & 0  \\  0 & _{-2}Y  \end{bmatrix}}
\newcommand{\ymatc}{\begin{bmatrix} _{+2}Y^* & 0  \\  0 & _{-2}Y^*  \end{bmatrix}}
\newcommand{\yzmat}{\begin{bmatrix} _{0}Y & 0  \\  0 & _{0}Y  \end{bmatrix}}
\newcommand{\yzmatc}{\begin{bmatrix} _{0}Y^* & 0  \\  0 & _{0}Y^*  \end{bmatrix}}
\newcommand{\ope}{\hat{O}_{E}}
\newcommand{\opb}{\hat{O}_{B}}
\newcommand{\bmat}{\begin{bmatrix}}
\newcommand{\emat}{\end{bmatrix}}
\newcommand{\mdi}{\mathcal{D}/\mathcal{I}}
\newcommand{\mm}{\mathcal{M}}
\newcommand{\md}{\mathcal{D}}
\newcommand{\mi}{\mathcal{I}}
\def\vp#1{$\bar{P}_{#1}$ }
\def\vs{$\bar{S}~$}
\def\eq#1{{Eq.~(\ref{#1})}}
\def\sec#1{{Sec.~\ref{#1}}}
\def\fig#1{{Fig.~\ref{#1}}}
\def\app#1{{Appendix~\ref{#1}}}
\def\ymat#1{\begin{bmatrix} _{+2}Y_{#1} & 0  \\  0 & _{-2}Y_{#1}  \end{bmatrix}}
\def\ymatc#1{\begin{bmatrix} _{+2}Y^{\dag}_{#1} & 0  \\  0 & _{-2}Y^{\dag}_{#1}  \end{bmatrix}}
\def\yzmat#1{\begin{bmatrix} _{0}Y_{#1} & 0  \\  0 & _{0}Y_{#1}  \end{bmatrix}}
\def\yzmatc#1{\begin{bmatrix} _{0}Y^*_{#1} & 0  \\  0 & _{0}Y^*_{#1}  \end{bmatrix}}

@@

\abstract{
  \noindent 
  \revisit{  We derive full-sky, real-space operators that convert between polarization Stokes $Q$/$U$ parameters to the coordinate-independent, scalar $E$/$B$ modes that are widely used in Cosmic Microwave Background and cosmic shear analysis.  The covolution kernels split naturally into angular and radial parts, and we show explicitly how the spatial extent of the convolution kernel depends on the targeted band-limit.  We show that an arbitrary radial dependence can produce $E$/$B$-like maps and that these are simply filtered versions of the standard $E$/$B$ maps.  This allows us to compute $E$/$B$ maps in real space with a compactly-supported kernel, an approach that can guarantee the avoidance of known foreground regions and can be employed in a massively parallel scheme at high-resolution.  We can compute power spectra using standard techniques, and recover the power spectrum of the sky with a simple window function.  The spatial structure of the real space operator provides some intuition for the $E/B$ structure of polarized, filamentary galactic foregrounds.
    %
     We cast the standard CMB polarization analysis operators in a matrix-vector notation which facilitates the derivations and shows that the kernels relate directly to spin-0 $Y_{\ell 2}$ spherical harmonic functions.  This new notation also allows us to derive real space operators which decompose the measured Stokes parameters into their even and odd-parity parts, without ever evaluating the scalar $E$/$B$ fields themselves.  This paper is the first part in a series of papers that explore real-space computation of polarization modes and their applications.
  } 
  \\
  \\
%  We derive fully sky real space operators to translate Stokes Q \& U parameters to scalars $E$ \& $B$ and vice versa. We explicitly show that these local real space operator are fully characterized by the spin-0 $Y_{\ell 2}$ spherical harmonic functions. These real space operators make transparent the association of radial and tangential patterns of polarization with the E modes and that of  the clockwise and anti-clockwise pinwheel patterns of polarization with B-modes. We cast the standard CMB polarization analysis operators in a matrix-vector notation which elucidate these derivations. Using this new notation also allows us to derive real space operators which decompose the measured Stokes parameters into those corresponding to E-modes and B-modes respectively, without ever evaluating the scalar fields themselves. We use these analytical derivations us to quantify the non-local nature of the relation between the two representations of the polarization field. We present a prescription for generalizing these operators, which allow the non-local behavior of these operators to be a tunable parameter. We derive constraints on these generalized operators to ensure reliable recovery of the standard $E$ and $B$ fields and the corresponding power spectra. \revisit{This is part 1 of a series of papers where we have derived the real space operators. In the following papers we will explore some of their applications.} \comment{Needs better blending.}
}

\maketitle
\begingroup
\let\clearpage\relax
\include{./sections/introduction}
\include{./sections/real_space_formalism}
\include{./sections/kernel_discussion}
\include{./sections/generalizedeb}
\include{./sections/discussion}
\include{./sections/acknowledgments}
\include{./sections/appendix}
\endgroup

\bibliographystyle{JHEP}
\bibliography{ref}

>>>>

\usepackage[latin1]{inputenc}
\usepackage{amsmath, amsfonts, amssymb, hyperref, graphicx, multirow, subfigure, bbold}
\hypersetup{colorlinks=True,citecolor=blue}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage[normalem]{ulem}
\usepackage{caption}
\usepackage{array}
\usepackage{aas_macros}

\renewcommand{\thesubfigure}{(\roman{subfigure})}
%\renewcommand{\thesubfigure}{}

\newcommand{\na}{New Astronomy}
%\newcommand{\aap}{Astronomy and Astrophysics}
\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}
\newcommand{\beqry}{\begin{eqnarray}}
\newcommand{\eeqry}{\end{eqnarray}}
\newcommand{\beqrys}{\begin{subequations}\begin{eqnarray}}
\newcommand{\eeqrys}{\end{eqnarray}\end{subequations}}
\newcommand{\fqu}{\begin{bmatrix} Q \\ U \end{bmatrix}}
\newcommand{\xpm}{\begin{bmatrix} _{+2}X \\ _{-2}X \end{bmatrix}}
\newcommand{\equ}{\begin{bmatrix} Q \\ U \end{bmatrix}_{E}}
\newcommand{\bqu}{\begin{bmatrix} Q \\ U \end{bmatrix}_{B}}
\newcommand{\eblm}{\begin{bmatrix} a^E \\ a^{B} \end{bmatrix}}
\newcommand{\qutox}{\begin{bmatrix} \mathbb{1} & i\mathbb{1}  \\  \mathbb{1} & -i\mathbb{1}   \end{bmatrix}}
\newcommand{\qutoxd}{\begin{bmatrix} \mathbb{1} & \mathbb{1}  \\  -i\mathbb{1} & i\mathbb{1}   \end{bmatrix}}
\newcommand{\ymat}{\begin{bmatrix} _{+2}Y & 0  \\  0 & _{-2}Y  \end{bmatrix}}
\newcommand{\ymatc}{\begin{bmatrix} _{+2}Y^* & 0  \\  0 & _{-2}Y^*  \end{bmatrix}}
\newcommand{\yzmat}{\begin{bmatrix} _{0}Y & 0  \\  0 & _{0}Y  \end{bmatrix}}
\newcommand{\yzmatc}{\begin{bmatrix} _{0}Y^* & 0  \\  0 & _{0}Y^*  \end{bmatrix}}
\newcommand{\ope}{\hat{O}_{E}}
\newcommand{\opb}{\hat{O}_{B}}
\newcommand{\bmat}{\begin{bmatrix}}
\newcommand{\emat}{\end{bmatrix}}
\newcommand{\mdi}{\mathcal{D}/\mathcal{I}}
\newcommand{\mm}{\mathcal{M}}
\newcommand{\md}{\mathcal{D}}
\newcommand{\mi}{\mathcal{I}}
\def\vp#1{$\bar{P}_{#1}$ }
\def\vs{$\bar{S}~$}
\def\eq#1{{Eq.~(\ref{#1})}}
\def\sec#1{{Sec.~\ref{#1}}}
\def\fig#1{{Fig.~\ref{#1}}}
\def\app#1{{Appendix~\ref{#1}}}
\def\ymat#1{\begin{bmatrix} _{+2}Y_{#1} & 0  \\  0 & _{-2}Y_{#1}  \end{bmatrix}}
\def\ymatc#1{\begin{bmatrix} _{+2}Y^{\dag}_{#1} & 0  \\  0 & _{-2}Y^{\dag}_{#1}  \end{bmatrix}}
\def\yzmat#1{\begin{bmatrix} _{0}Y_{#1} & 0  \\  0 & _{0}Y_{#1}  \end{bmatrix}}
\def\yzmatc#1{\begin{bmatrix} _{0}Y^*_{#1} & 0  \\  0 & _{0}Y^*_{#1}  \end{bmatrix}}

@@

\abstract{
  \noindent 
  \revisit{We derive full-sky, real-space operators that convert between polarization Stokes $Q/U$ parameters to the coordinate-independent scalar $E$/$B$ modes that are widely used in Cosmic Microwave Background(CMB) and cosmic shear analysis.  The covolution kernels split naturally into angular and radial parts and we show explicitly how the radial extent of the convolution kernel depends on the targeted band-limit.  We show that an arbitrary radial dependence can produce $E/B$-like maps and that these are simply filtered versions of the standard $E/B$ maps.  This allows us to compute $E/B$ maps in real space with a compactly-supported kernel, an approach that can guarantee the avoidance of known foreground regions and can be employed in a massively parallel scheme at high-resolution.  We can recover the standard power spectrum of the polarized CMB sky by correcting the power spectra of these maps with a simple window function.  The spatial structure of the real space operator provides great intuition for the $E/B$ structure of polarized, filamentary galactic foregrounds and we predict non-zero B-mode signature that is expected from finite filaments in the sky.
    %
     We cast the standard CMB polarization analysis operators in a matrix-vector notation which facilitates the derivations and shows that the kernels relate directly to spin-0 $Y_{\ell 2}$ spherical harmonic functions.  This new notation also allows us to derive real space operators which decompose the measured Stokes parameters into their even and odd-parity parts, without ever evaluating the scalar $E$/$B$ fields themselves.  This paper is the first part in a series of papers that explore real-space computation of polarization modes and their applications.
  } 
  \\
  \\
%  We derive fully sky real space operators to translate Stokes Q \& U parameters to scalars $E$ \& $B$ and vice versa. We explicitly show that these local real space operator are fully characterized by the spin-0 $Y_{\ell 2}$ spherical harmonic functions. These real space operators make transparent the association of radial and tangential patterns of polarization with the E modes and that of  the clockwise and anti-clockwise pinwheel patterns of polarization with B-modes. We cast the standard CMB polarization analysis operators in a matrix-vector notation which elucidate these derivations. Using this new notation also allows us to derive real space operators which decompose the measured Stokes parameters into those corresponding to E-modes and B-modes respectively, without ever evaluating the scalar fields themselves. We use these analytical derivations us to quantify the non-local nature of the relation between the two representations of the polarization field. We present a prescription for generalizing these operators, which allow the non-local behavior of these operators to be a tunable parameter. We derive constraints on these generalized operators to ensure reliable recovery of the standard $E$ and $B$ fields and the corresponding power spectra. \revisit{This is part 1 of a series of papers where we have derived the real space operators. In the following papers we will explore some of their applications.} \comment{Needs better blending.}
}

\maketitle
\begingroup
\let\clearpage\relax
\include{./sections/introduction}
\include{./sections/real_space_formalism}
\include{./sections/kernel_discussion}
\include{./sections/generalizedeb}
\input{./sections/kernel_paintbrush.tex}
\include{./sections/discussion}
\include{./sections/acknowledgments}
\include{./sections/appendix}
\endgroup

\bibliographystyle{JHEP}
\bibliography{ref,refkmh}

### 2018-06-11 articles/formalism/sections/discussion.tex
significant updates to discussions and polarized filaments sections

In this article we have cast the standard CMB polarization analysis operations in a vector matrix notation. Using this concise notation we derive the real space operators that translate the Stokes vector \vp{} to  the vector of scalars \vs and vice versa. We explicitly demonstrated that this real space operation can be simply interpreted as a convolution over the complex field $[Q - i U]$ (or $[E + iB]$) with an effective complex beam which is fully expressed in terms of the $Y_{\ell 2}$ spherical harmonic functions. We also use this vector matrix notation to derive real space operators which allow the direct decomposition of the full Stokes vector \vp{} into the vector \vp{E} and \vp{B} that correspond to the respective scalar modes.

Given the effective beam interpretation of these real space operators we derive the harmonic coefficients of these effective beams at the north galactic pole. Using these harmonic coefficients we provide a prescription for computing the convolution kernels at any position on the sphere using the standard Healpix built in functions. The procedure is equivalent to parallel transporting the beam at the north pole to any desired location on the sphere. We implement the prescription to compute the  kernel at different location on the sphere and provide simple explanations in terms of Euler angles for the observed variations.

These real space convolution kernels provide a spatially intuitive way of understanding the construction of the scalar modes. We explicitly show that the kernels separates into an band limit independent azimuthal operation around any given direction which is primarily responsible for requisite decomposition, while the band limit dependent radial weights can be interpreted as some isotropic smoothing operation. These radial weights primarily determine the non-local dependence of the construction of the respective fields at any location on the Stoke field. We define the parameter $\beta_0$ as a means to characterize the non-locality and show that $\beta_0$ scales $\propto \ell_{\rm max}^{-1}$. We show that this non-locality parameter also characterized the non-locality of the $\bar{O}_{E/B}$ operators.

Finally we present the generalized real space operators $\bar{O}'$, which are derived by allowing the radial function to vary from its default form. We derive constraints on the modifications to these radial function by demanding the inverse operator to be well defined. We argue that these modifications to the radial kernel can be be interpreted as a some smoothing smoothing operation on the scalar fields with a circularly symmetric instrument beam. We also show that as long as these radial function are invertible, the standard spectra can always be recovered from these modified $E'$ \& $B'$ maps. The main advantage of modifying these radial function is the ability to generate more locally defined $E$ and $B$ mode maps. This could potentially be useful in reducing foreground contamination on large angular scales in a full sky $E/B$ analysis. Also defining more locally constructed scalar fields $E$ \& $B$ can be used to circumvent the power leakage nuisance. We explore and demonstrate the working of these ideas in the next paper in this series. 
%\revisit{The discussion till now gives the impression that using the localized convolution kernels is no different from from using the default kernel and altering the spherical harmonic coefficients of expansion of the relavant fields by appropriately operating on them with the  effective beam functions $g_{\ell}$. To appreciate the difference between these two, it is important to realize that in general one can make a choice of a radial function which may not have a band limited description. In such a case these two method of evaluating the relevant fields is not identical. An example of this claim is depicted in \fig{fig:example_gbeta}.\\
%Another important thing to realize is that the harmonic coefficients derived from default full sky operations get some contributions from different portions of sky. For instance evaluating the E and B fields in the vicinity of the poles is are prone to receiving significant contributions from strong foregrounds near the equator. Correcting the harmonic coefficients of expansion with the effective beam function does not cancel these non-local contribution. On the contrary by performing the convolution with the localized real space kernels, the regions which contribute to the local field evaluations are predetermined by the choice of the radial function.}

>>>>

We introduced a vector-matrix notation which allows for concise book keeping of all the standard operations involved in the analysis of CMB polarization maps. This notation simplifies the derivation of the real space operators. We presented a first derivation of the real space operators on the sphere that transforms the Stokes vector \vp{} to into a vector of scalars \vs{} and vice versa. We also presented real space operators that directly decompose the full Stokes vector \vp{} into vector \vp{E} and \vp{B} that correspond to the respective scalar modes.

These real space operators provide a spatially intuitive way of understanding the different decompositions of the Stokes vector. We explicitly  demonstrated that all the real space operators have the common characteristic  of being separable into a azimuthal part that is band limit independent and radial part that incorporates all the band limit dependence. The azimuthal part of the operator is primarily responsible for the requisite spinorial decomposition, while the radial weights determine the non-local dependence of the construction of the resultant fields on the original fields. Using the self similarity property of the radial functions we define a non-locality parameter and argued that it is proportional to $\ell_{\rm max}^{-1}$. The scalar decomposition of the Stokes parameters being of primary interest,  we empirically derived the non-locality parameter $\beta_0$ for this operation and showed that the proportionality constant $\ell_{0}=22$ provided a good prediction for the angular distance at which the ${}_{\mm}f$ falls below 1\% of its maxima as seen in \fig{fig:rad_ker_decay}.

Our careful study of the real space operators reveals the dual interpretation of the operators as either the Green's function or a convolving beam depending on whether the kernels are expressed in terms of the forward rotation Euler angles or the inverse rotation Euler angles. While the convolution interpretation is a familiar one, the Green's function interpretation is a new way of looking at this operation. This new interpretation allows us to think of ${}_{+2}X$ as some spin-2 charge which radiates out a complex spin-0 scalar field $E+iB$. The resultant complex scalar maps can be then understood as arising from superposition of the radiating field emanating from all the spin charges on the sphere. The $E/B$ mode maps are merely the real and imaginary parts of this resultant complex scalar field.

\revisit{A similar equation for real space $E$ \& $B$ operators was derived in \cite{Zaldarriaga2001a}, however those results were derived for the flat sky case and did not explicitly derive the radial kernel.} \comment{A discussion on this should be in the conclusions.}

Understanding the non-locality of the real space operators, allowed us to generalize the real space operators $\bar{O}'$ and its inverse ${\bar O}'^{-1}$, which transform between the spin-2 and spin-0 representations of the CMB polarization. We presented a method of systematically generalizing these operators by introducing the harmonic space operator $\tilde{\mathcal{G}}$ which in effect causes the radial function to vary from its default form. We show that these modifications to the radial kernel can be interpreted as a smoothing operation on the scalar fields with a circularly symmetric instrument beam. We argue that one can choose any arbitrary form for $\tilde{\mathcal{G}}$ as long as one adheres to the constraint that $\tilde{\mathcal{G}}^{-1}$ and $\tilde{\mathcal{G}}^{-2}$ are well defined operators.  These constraints are necessary to ensure that all the standard CMB spectra are recoverable from the modified $E'$/$B'$ maps emerging from the modified operators on the Stokes parameter maps. Noting that that the standard spin raising $\eth^2$ and spin lowering $\bar{\eth}^2$ operators are special cases of these generalized operators, we presented a band limited representation of these operators.

\revisit{Discussion on results from polarized filaments filaments, compare and contrast with those in \cite{Zaldarriaga2001a}.}

There are several advantages and alternative analysis routes that emerge from employing these real space operators for analysis of CMB polarization maps. To begin with, one does not need to evaluate the spin harmonic functions as the real space operators only rely on computing the Euler angles which can be done on the fly and the $P_{\ell}^{2}$ functions need to be tabulated only once for analysis at some predetermined resolution. Note that this real space analysis method trivially generalizes to analyzing maps of arbitrary spin. Especially given the Green's function interpretation of these operators, their implementation is trivially parallelizable, since the $E/B$ contribution from the Stokes charge in each pixel can be evaluated independently. Also note that the evaluation of contribution from charges in the masked portion of the sky can be easily omitted, unlike in the harmonic analysis. Given the Green's function interpretation of the  real space operators and their locality, suggests the possibility of converting the Stokes parameter time ordered data to scalar $E/B$ time ordered data given the pointing model for the observations, which can then be subsequently converted to maps of $E/B$ modes using standard map making techniques. Since the real space operators offer the ability of tuning the locality of $E/B$ maps, this can be potentially exploited to reduce foreground contamination from distant parts of the sky. For instance the real space operators can be defined such that the locality of their radial kernels is varied on different portions of the sky, dictated by say the foreground morphology, resulting in some scalar $E'/B'$ maps.  While this idea seems promising, the usefulness of this implementation will depend on whether the standard $E/B$ mode spectra are recoverable from scalar maps derived in this fashion. Note that it would be fairly non-trivial to think of such implementations using conventional harmonic space methods.

Results demonstrating the equivalence of these real space methods to the conventional harmonic space methods and from explorations of some of the ideas discussed above will be presented in the second part of this series of papers.

%\revisit{The discussion till now gives the impression that using the localized convolution kernels is no different from from using the default kernel and altering the spherical harmonic coefficients of expansion of the relavant fields by appropriately operating on them with the  effective beam functions $g_{\ell}$. To appreciate the difference between these two, it is important to realize that in general one can make a choice of a radial function which may not have a band limited description. In such a case these two method of evaluating the relevant fields is not identical. An example of this claim is depicted in \fig{fig:example_gbeta}.\\
%Another important thing to realize is that the harmonic coefficients derived from default full sky operations get some contributions from different portions of sky. For instance evaluating the E and B fields in the vicinity of the poles is are prone to receiving significant contributions from strong foregrounds near the equator. Correcting the harmonic coefficients of expansion with the effective beam function does not cancel these non-local contribution. On the contrary by performing the convolution with the localized real space kernels, the regions which contribute to the local field evaluations are predetermined by the choice of the radial function.}

### 2018-06-11 articles/formalism/sections/generalizedeb.tex
significant updates to discussions and polarized filaments sections

%--------------------------------------------------------
\subsubsection{Relation to the spin raising $\eth^2$ and lowering $\bar{\eth}^2$ operators}
Recall that on operating twice with the spin lowering operator on the Stokes charge ${}_{-2}X$ results in filtered version of $E/B$ maps as in \eq{eq:ebdef}. Now note that it is possible to construct a modified real space operator by choosing the harmonic space function to be $g_{\ell} = \sqrt{\frac{(\ell+2)!}{(\ell-2)!}}$, resulting in similarly filtered $E/B$ maps as follows:
%
\beq
[\mathcal{E} + i \mathcal{B}](\hat{n}_e)=- \Delta \Omega\sum_{q=1}^{N_{\rm pix}} \Bigg\lbrace  \left[  \sum_{\ell=\ell_{\rm min}}^{\ell_{\rm max}} \frac{2 \ell+1}{4 \pi} P_{\ell}^2(\beta_{qe}) \right] e^{-i2\alpha_{eq}} {}_{+2}X(\hat{n}_{q}) \Bigg\rbrace \,. \label{eq:bl_ebdef_lower}
\eeq
%
Comparing \eq{eq:ebdef_lower} to \eq{eq:bl_ebdef_lower} makes apparent the following mapping:
%
\beq
\bar{\eth}^2 \equiv \Delta \Omega \sum_{q=1}^{N_{\rm pix}} \left[ \sum_{\ell=\ell_{\rm min}}^{\ell_{\rm max}} \frac{2 \ell+1}{4 \pi} P_{\ell}^2(\beta_{qe}) \right] e^{-i2\alpha_{eq}}\,.
\eeq
%
The operator $\bar{\eth}^2$ is composed of derivative operations and hence has an implicit bi-positional dependence for all numerical purposes which is made explicit in the band limited version of the operator.  The band limited version of the spin raising operator $\eth^2$ is derived by merely taking the conjugate of the above equation.
%--------------------------------------------------------

>>>>

%--------------------------------------------------------
\subsubsection{Relation to the spin raising $\eth^2$ and lowering $\bar{\eth}^2$ operators}
Recall that on operating twice with the spin lowering operator on the Stokes charge ${}_{+2}X$ results in filtered version of $E/B$ maps as in \eq{eq:ebdef}. Now note that it is possible to construct a modified real space operator by choosing the harmonic space function to be $g_{\ell} = \sqrt{\frac{(\ell+2)!}{(\ell-2)!}}$, resulting in similarly filtered $E/B$ maps as follows:
%
\beq 
\label{eq:bl_ebdef_lower}
[\mathcal{E} + i \mathcal{B}](\hat{n}_e)=- \Delta \Omega\sum_{q=1}^{N_{\rm pix}} \Bigg\lbrace  \left[  \sum_{\ell=\ell_{\rm min}}^{\ell_{\rm max}} \frac{2 \ell+1}{4 \pi} P_{\ell}^2(\beta_{qe}) \right] e^{-i2\alpha_{eq}} {}_{+2}X(\hat{n}_{q}) \Bigg\rbrace \,. 
\eeq
%
Comparing \eq{eq:ebdef_lower} to \eq{eq:bl_ebdef_lower} makes apparent the following mapping:
%
\beq
\bar{\eth}^2 \equiv \Delta \Omega \sum_{q=1}^{N_{\rm pix}} \left[ \sum_{\ell=\ell_{\rm min}}^{\ell_{\rm max}} \frac{2 \ell+1}{4 \pi} P_{\ell}^2(\beta_{qe}) \right] e^{-i2\alpha_{eq}}\,.
\eeq
%
The operator $\bar{\eth}^2$ is composed of derivative operations and hence has an implicit bi-positional dependence for all numerical purposes which is made explicit in the band limited version of the operator.  The band limited version of the spin raising operator $\eth^2$ is derived by merely taking the conjugate of the above equation.
%--------------------------------------------------------

### 2018-06-11 articles/formalism/sections/kernel_discussion.tex
significant updates to discussions and polarized filaments sections

To quantify the non-locality of the scalar modes $E/B$, we can define a characteristic angular radius of the region from which the kernels get most of their contribution.   We define a non-locality parameter $\beta_{0}$ as the angle beyond which the function ${_{\mm}f}(\beta,\ell_{\rm min}=2,\ell_{\rm max})$ consistently falls below 1\% of its maximum.
%For $\ell_{\rm max}=24$, the maximum multipole accessible on a Nside=8 Healpix map, the non-locality parameter $\beta_0=180^{\circ}$ as the radial function never falls monotonously below 1\% of its global maxima as seen in \fig{}. Using this fact and the self similar property of the radial functions, we define the following empirical relation: $\beta_o= {\rm min}(180,180 \frac{24}{\ell_{\rm max}})$, as a means of estimating the non-locality parameter given the maximum multipole $\ell_{\rm max}$ accessible for analysis.
\revisit{The empirical relation:
\beq
\beta_0= {\rm min}\left(180,180 \frac{\ell_{0}}{\ell_{\rm max}} \right) \,,
\eeq
with $\ell_{0}=22$ provides a reasonable estimate of this transition point for ${}_{\mm}f$ as seen in \fig{fig:rad_ker_decay}. Setting $\ell_{0}=10$ and $\ell_{0}=32$ provides the transition points for the functions ${}_{\mi}f$ and  ${}_{\md}f$ respectively.}

>>>>

To quantify the non-locality of the scalar modes $E/B$, we can define a characteristic angular radius of the region from which the kernels get most of their contribution.   We define a non-locality parameter $\beta_{0}$ as the angle beyond which the function ${_{\mm}f}(\beta,\ell_{\rm min}=2,\ell_{\rm max})$ consistently falls below 1\% of its maximum.
%For $\ell_{\rm max}=24$, the maximum multipole accessible on a Nside=8 Healpix map, the non-locality parameter $\beta_0=180^{\circ}$ as the radial function never falls monotonously below 1\% of its global maxima as seen in \fig{}. Using this fact and the self similar property of the radial functions, we define the following empirical relation: $\beta_o= {\rm min}(180,180 \frac{24}{\ell_{\rm max}})$, as a means of estimating the non-locality parameter given the maximum multipole $\ell_{\rm max}$ accessible for analysis.
\revisit{The empirical relation:
\beq
\beta_0= {\rm min}\left(180,180 \frac{\ell_{0}}{\ell_{\rm max}} \right) \,,
\eeq
with $\ell_{0}=22$ provides a reasonable estimate of this transition point for ${}_{\mm}f$ as seen in \fig{fig:rad_ker_decay}. Setting $\ell_{0}=10$ and $\ell_{0}=32$ predicts the transition points for the functions ${}_{\mi}f$ and  ${}_{\md}f$ respectively.} \comment{How sensitive is this transition point to $\ell{\rm min}$?}

\comment{Compare the fitted slope to the one in the Zaldariaga paper.}

### 2018-06-11 articles/formalism/sections/kernel_paintbrush.tex
Added colorbar
significant updates to discussions and polarized filaments sections

\paragraph{Polarization signature of magnetized filaments}
% 
\begin{figure}[t]
\includegraphics[width=0.5\columnwidth]{line.pdf}
\includegraphics[width=0.5\columnwidth]{spiral.pdf}
\caption{ The polarization signals of toy filament structures. In a filament organized perfectly along a magnetic field line, the polarization will be perpendicular to the filament direction.  The $E/B$ modes of filaments are in some ways easier to think about than the Stokes parameters. Left panels: in a straight filament, the E-mode is positive along the filament and at the ends, but negative along the sides.  B-modes are only non-zero at the ends.  Right panels: in a curved filament, the E-mode is again positive along the filament.  Outside the filament, the $E$-mode is more negative on the interior of the curve than the exterior.  The $B$-modes are again non-zero only at the ends, and are akin to the straight filament case. In all images, the longitude angle increases to the left (East in sky convention).\comment{Colorbar on plots to give a sense of B amplitude as compared to E.}}
\label{fig:polfilaments}
\end{figure}
%

@@

The non-zero $B$ result is somewhat surprising given that the polarization pattern is symmetric to both horizontal and vertical reflections through the filament center.  However, unlike a complete ring, this filament is not a configuration with a definite parity.  Because the scalar polarization descriptions are coordinate independent, the $E/B$ patterns do not depend on the orientation of the filament.  A filament inclined at $45^\circ$ will have a similar $E/B$ pattern, but different reflection symmetries.

%\revisit{A stacking analysis of Planck data \citep{2016A&A...586A.141P} sees $E>0$ along filaments (selected from intensity data), but no $B$-mode signal is above the noise.  We predict that it should be there in higher fidelity data.}
\revisit{A stacking analysis of Planck data \citep{2016A&A...586A.141P} sees $E>0$ along filaments (selected from intensity data), but claim no $B$-mode signal. We predict that a B-mode signal from filaments is present. A naive filament stacking analysis is only ideal for making E-mode detections. For detecting B-modes from filaments requires a more careful edge stacked filaments analysis. While its detectability from Planck requires more study, it should definitely be detectable in higher fidelity data.}

The intuition from the real-space kernels holds also when we distort the shape of the filament.  If the filament were bent around into a circle \comment{Not for an ellipse right? Maybe constant curvature argument is more general and more precise}, the positive and negative parts of the $B$ pattern will cancel, and we are left with a hoop of pure $E$ pattern.  The same general description holds for a spiral-shaped filament, which can can be viewed as distortion of the straight filament.  The filament is highlighted by positive $E>0$.  The $E$-pattern is more negative on the interior of a curve than on the exterior, and the concentric rings of filamentary structure make an increasingly negative $E$ value inside.  The $B$-pattern is again concentrated at the ends of the filament in an oriented pair of positive/negative fluctuations.

>>>>

\section{Understanding polarization signatures of magnetized filaments}
% 
\begin{figure}[t]
\includegraphics[width=0.5\columnwidth]{line.pdf}
\includegraphics[width=0.5\columnwidth]{spiral.pdf}
\includegraphics[width=1.0\columnwidth]{colorbar.pdf}
\caption{ The polarization signals of toy filament structures. In a filament organized perfectly along a magnetic field line, the polarization will be perpendicular to the filament direction.  The $E/B$ modes of filaments are in some ways easier to think about than the Stokes parameters. Left panels: in a straight filament, the E-mode is positive along the filament and at the ends, but negative along the sides.  B-modes are only non-zero at the ends.  Right panels: in a curved filament, the E-mode is again positive along the filament.  Outside the filament, the $E$-mode is more negative on the interior of the curve than the exterior.  The $B$-modes are again non-zero only at the ends, and are akin to the straight filament case. In all images, the longitude angle increases to the left (East in sky convention).  All plot are on a common, arbitrary color scale.}
\label{fig:polfilaments}
\end{figure}
%

@@

The non-zero $B$ result is somewhat surprising given that the polarization pattern is symmetric to both horizontal and vertical reflections through the filament center.  However, unlike a circular ring, this filament is not a configuration with a definite parity.  Because the scalar polarization descriptions are coordinate independent, the $E/B$ patterns do not depend on the orientation of the filament.  A filament inclined at $45^\circ$ will have a similar $E/B$ pattern, but different reflection symmetries.

%\revisit{A stacking analysis of Planck data \citep{2016A&A...586A.141P} sees $E>0$ along filaments (selected from intensity data), but no $B$-mode signal is above the noise.  We predict that it should be there in higher fidelity data.}
\revisit{A stacking analysis of Planck data \citep{2016A&A...586A.141P} sees $E>0$ along filaments (selected from intensity data), but claim no $B$-mode signal. We predict that a B-mode signal from filaments should be present. A naive filament stacking analysis is only suited for making E-mode detections. Detecting B-modes from filaments requires a more careful edge stacked filaments analysis. While the detectability of B-mode signature from filaments from Planck data calls for a more careful assessment, it should definitely be detectable in higher fidelity data.}

The intuition from the real-space kernels holds also when we distort the shape of the filament.  If the filament were bent around into a circle, the positive and negative parts of the $B$ pattern will cancel, and we are left with a hoop of pure $E$ pattern (note that this cancellation would not happen for an ellipse or any loop of non-contant radius of curvature).  The same general description holds for a spiral-shaped filament, which can can be viewed as distortion of the straight filament.  The filament is highlighted by positive $E>0$.  The $E$-pattern is more negative on the interior of a curve than on the exterior, and the concentric rings of filamentary structure make an increasingly negative $E$ value inside.  The $B$-pattern is again concentrated at the ends of the filament in an oriented pair of positive/negative fluctuations. \comment{Add that the spiral has a constant radius of curvature and thats the reason you dont see B-mode signal in intermediate regions of the spiral!}

\revisit{Note that we have made the simplifying assumption that the polarization headless vectors are perpendicular to the tangent to the filament at any given point and this results in dominantly generating an E-mode signal while the B-mode signal is present only near the edges. By similar arguments it can be seen that that polarization pattern will be dominated by B-modes if the polarization vector is always at $+45$ (or $-45$) degrees to the tangent to the filament while the E-mode signal would only be present at the edges. Therefore a careful study of the $E/B$ mode power in a filament can provide insights into the orientation of the magnetic field with respect to the length of the filament, assuming the polarization vectors (headless) are always perpendicular to the magnetic field vector.}
